### 2017-01-22 docs/shared_task_paper.tex
added an outline for the shared task paper

\documentclass[11pt]{article}
\usepackage{eacl2017}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

% \eaclfinalcopy % Uncomment this line for the final submission
%\def\eaclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Story Cloze task -- UW System}

\author{Roy Schwartz\\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And	
  Maarten Sap \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Taking part in the LSDSem `17 shared task, we describe our system.

\end{abstract}

\section{Intro}

\section{System Overview}

\section{Features}

\section{Results}

\section{Discussion \& Conclusion}

\bibliography{acl2017}
\bibliographystyle{eacl2017}

\end{document}