### 2017-01-29 docs/shared_task_paper.tex
outlined some of the shared task stuff

\documentclass[11pt]{article}
\usepackage{eacl2017}
\usepackage{times}
\usepackage{url}
\usepackage{latexsym}

% \eaclfinalcopy % Uncomment this line for the final submission
%\def\eaclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Story Cloze task -- UW System}

\author{Roy Schwartz\\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And	
  Maarten Sap \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
Taking part in the LSDSem `17 shared task, we describe our system.

\end{abstract}

\section{Intro}
% Describe the overall landscape of the problem
% Mention the specific purpose of the shared task (i.e. from the website)
% Last paragraph should be a quick description of our system.
In this paper, we describe the system submission from the University of Washington NLP group (UW NLP).
We implement a simple classifier that uses surface features as well as language model features. Using the Shared Task test set, we achieve $75.1\%$ accuracy.

\section{System Overview}
%% Explain that we have two sets of features, style and language.
%% Maybe mention here that we're doing training on validation set ?

\section{Features}
\subsection{Stylistic Features}
%% take stuff from the paper

\subsection{Language Model Features}
%% take the RLM paragraph from the paper

\section{Results \& Discussion}
% Show results

% ablation study? Table 2 from the paper sort of?

\section{Conclusion}

% Our system makes use of knowledge brought to us by the constrained writing literature (?)
% talk about how we use state-of-the-art language modelling tools.
% We achieve $75.1\%$ which is ranked XXth in the ongoing codalab competition.

\bibliography{acl2017}
\bibliographystyle{eacl2017}

\end{document}